\documentclass[a4paper,twoside, 12pt]{memoir} %A4papir, to side, størrelse 12, type memoir 	
% For dansk opsætning med æ, ø og å, samt pænere orddeling.
\usepackage[utf8]{inputenc}	
\input{preamble}
\begin{document}
\include{frontpage}
\chapter*{Preface}
    This bachelor's thesis was written for the Department of Electrical and Computer Engineering, Aarhus University. It is part of the Computer Engineering study program, and was written in the spring semester of 2024.\\
    Special thanks goes to my supervisor Hugo Daniel Macedo for suggesting the initial problem domain, connecting me with relevant contacts, and supervising this thesis.\\
    \textbf{All source files associated with this thesis are found at:} \url{https://github.com/asgersong/BSc}
    \vspace{1.5cm}

    \begin{flushright}
        \textit{Asger Poulsen, \today}
    \end{flushright}
\chapter*{Abstract}
This bachelor's thesis delves into the integration of linear algebra concepts in the development and optimization of Large Language Models (LLMs) for software engineering applications.
The thesis presents the theoretical underpinnings of LLMs, particularly focusing on the transformer architecture and its components. Furthermore, it explores practical applications through the development of a Retrieval-Augmented Generation (RAG) chatbot prototype, designed to assist students by providing contextually accurate and detailed responses. 
The chatbot leverages state-of-the-art models and integrates user-uploaded documents to enhance interaction quality. Additionally, a case study on the low-rank approximation of the BART model assesses the effectiveness of reducing computational complexity and storage requirements while maintaining performance on summarization tasks. 
The results, evaluated through ROUGE scores, underscore that low-rank approximation can significantly optimize model efficiency, but also highlight the trade-offs between performance and resource optimization. This work contributes to bridging the gap between theoretical linear algebra concepts and their practical applications in modern natural language processing tasks (NLP), offering insights for future research and development in the field.

\emph{\textbf{Keywords:}} Large Language Models, Linear Algebra, Retrieval-Augmented Generation, Chatbot, BART, Low-Rank Approximation, Summarization, ROUGE Scores
\newpage
\chapter*{Resume}
Denne bacheloropgave dykker ned i integrationen af lineære algebra-koncepter i udviklingen og optimeringen af store sprogmodeller (LLMs) til anvendelser inden for software engineering. 
Opgaven præsenterer de teoretiske fundamenter for LLM'er, med særlig fokus på transformer-arkitekturen og dens komponenter. 
Derudover udforskes praktiske anvendelser gennem udviklingen af en prototype på en Retrieval-Augmented Generation (RAG) chatbot, designet til at assistere studerende ved at levere kontekstuelt præcise og detaljerede svar. 
Chatbotten udnytter avancerede modeller og integrerer bruger-uploadede dokumenter for at forbedre interaktionskvaliteten. Yderligere vurderes en case-studie om lav-rangs-approksimering af BART-modellen, der undersøger effektiviteten af at reducere beregningskompleksitet og lagerkrav, samtidig med at ydeevnen på summariseringsopgaver opretholdes. 
Resultaterne, evalueret gennem ROUGE-scores, understreger, at lav-rangs-approksimering kan optimere modeleffektivitet markant, men fremhæver også afvejningerne mellem ydeevne og ressourceoptimering. Dette arbejde bidrager til at bygge bro mellem teoretiske lineære algebra-koncepter og deres praktiske anvendelser i moderne opgaver inden for naturlig sprogbehandling (NLP), og tilbyder indsigt til fremtidig forskning og udvikling på området.

\newpage
\tableofcontents*

%% Here is where you can include your chapters
\include{Introduction}
\include{TheoreticalFoundations}
\include{LiteratureReview}
\include{Methodology}
\include{Implementation}
\include{EvaluationAndResults}
\include{Discussion}
\include{Conclusion}
%%
\printbibliography

% Appendices
\appendix
\include{Appendix-Education}
\include{Appendix-Interview}
\include{Appendix-BART}

\end{document}
